---
title       : Working Hours per week in different countries, years 1985 - 2007
subtitle    : Developing Data Products course project (Coursera)
author      : Andrey Budish
job         : 30th of April, 2017
framework   : io2012    # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---


## Project Description

* The main goal of this project was to create Shiny application and provide a reproducible pitch for it in a form of Slidify presentation.  

* I have decided to use open data from [gapminder] (https://www.gapminder.org/data/), specifically [working hours per week data set] (https://docs.google.com/spreadsheets/d/1WPjCgWnDnRY_3wdW9_5KuIDHS-Lz1fML8Dcn0QNGiSs/pub) provided by International Labor organization.  

* The idea behind the app is to compare two countries by working hours per week during a specific time period.  

* The app itself is available [here] (https://andrey-budish.shinyapps.io/workinghours/)  

* All source files, Shiny app related code and this presentation, are available at [https://github.com/abudish/workingHours] (https://github.com/abudish/workingHours)

---

## The Working Hours per Week App
* This app allows you to see how working hours per week has changed for one country and compare it to another country for the same time period

* To use the app one should choose two countries from available listboxes.  

* Then choose a range of years (available range is 1985-2007)

* The results are displayed in 3 tabs:
  > 1. Tab 'Plot': displayes time series plot for the selected countries and period
  > 2. Tab 'Selected data': selected data in a form of a data table with variables country, year and hours_per_week
  > 3. Tab 't-Test': shows  the result of Welch Two Sample t-test that checks if there is a difference between hours_per_week variable for the selected countries and period.

---

## Data set
Original data set is not 'tidy', so it needs to be transformed in the way that each row is a unique observation. Transformed data set contains 3 variables: country, year and hour_per_week. Total number of unique countries is 25, year range is from 1985 to 2007, number of observations is 575.

```{r}
library(reshape2); library(readxl)
df <- read_xlsx("Working hours per week (1985 - 2007).xlsx")
df <- melt(df, id="Working hours per week")
colnames(df) <- c("country", "year", "hours_per_week")
head(df, 3)
```

---

## Plot example
Below are the output plot, if user's input was: range - from 1991 to 2000, countries - Australia and Canada. 

```{r, echo=FALSE, fig.width=9, fig.height=6, fig.align='center'}
library(ggplot2); library(ggthemes)
year_range <- 1991:2000
selected_countries <- subset(df, (year %in% year_range) & (country=="Australia" | country=="Canada"))
ggplot(data=selected_countries,
       aes(x=year, y=hours_per_week, colour=country, group=country)) + 
  geom_line(size=1.5) +
  geom_point(size=2, shape=21, fill="red") +
  ggtitle("Working hours per week, Australia and Canada") +
  theme_hc() +
  scale_colour_hc() +
  theme(plot.title = element_text(hjust = 0.5, size = 22),
        text = element_text(size = 15)) +
  labs(y= "Working hours per week", x="Years")
```


